---
title: "MD 10 Demo"
format: html
editor: visual
embed-resources: true
---

```{r}
#| message: false

library(tidyverse)
library(broom)
library(moderndive)
library(ggplot2movies)
library(ggeffects)
theme_set(theme_light())
```

Has the relationship between movie length and average rating changed over time? I will compare films from 1974 and 2004 to see if this relationship has changed.

## prep and wrangling

```{r}
#| message: false

data(movies)
glimpse(movies)

features2 <- movies |>
  filter(year == 2004 | year == 1974) |> 
  filter(Short == 0 & 
           length < 240 &
           length >= 80 ) |> 
  mutate(year = factor(year))
```

# descriptives

```{r}
ggplot(features2,
       aes(x = length,
           fill = year)) +
  geom_density(alpha = .2,
               color = NA)
```

```{r}
ggplot(features2,
       aes(x = rating,
           fill = year)) +
  geom_density(alpha = .2,
               color = NA)
```

```{r}
ggplot(features2,
       aes(x = length,
           y = rating,
           color = year )) +
  geom_jitter(alpha = .2) +
  geom_smooth(method = "loess")
```

## possible models

```{r}
m1 <- lm(rating ~ length + year + Action + Animation + 
           Comedy + Drama + Documentary + Romance,
         data = features2)
tidy(m1)

ggpredict(m1,
          terms = c("length", "year")) |> 
  plot()
```

```{r}
m2 <- lm(rating ~ length + I(length^2) + year + 
           Action + Animation + Comedy + Drama + 
           Documentary + Romance,
         data = features2)
tidy(m2)

ggpredict(m2,
          terms = c("length", "year")) |> 
  plot()
```

```{r}
m3 <- lm(rating ~ length * year + 
           Action + Animation + Comedy + Drama + 
           Documentary + Romance,
         data = features2)
tidy(m3)

ggpredict(m3,
          terms = c("length", "year")) |> 
  plot()
```

```{r}
m4 <- lm(rating ~ (length + I(length^2)) * year + 
           Action + Animation + Comedy + Drama + 
           Documentary + Romance,
         data = features2)
tidy(m4)

ggpredict(m4,
          terms = c("length", "year")) |> 
  plot()
```

## compare models

```{r}
get_regression_summaries(m1)
get_regression_summaries(m2)
get_regression_summaries(m3)
get_regression_summaries(m4)

## the AIC (Akaike's Information Criterion) is AIC = N*log(RSS/N) + 2k
## choosing the lowest is generally a reasonable idea

AIC(m1,m2,m3,m4)
```

## diagnostics

```{r}

m3_points <- get_regression_points(m3)

## linearity

ggplot(m3_points,
       aes(x = rating_hat,
           y = residual)) +
  geom_jitter(alpha = .2) +
  geom_hline(yintercept = 0,
             color = "red")

## normality

ggplot(m3_points,
       aes(x = residual)) +
  geom_histogram(boundary = 0,
                 binwidth = .5,
                 color = "white")
```

## interpretation

```{r}
tidy(m3, conf.int = TRUE)

ggpredict(m3,
          terms = c("length", "year")) |> 
  plot()
```

Questions:
* What is the expected difference in rating between a 90-minute and 120-minute film in 2004? In 1974?
* Why do you think this has changed?
* Should studios just make longer movies to increase critical acclaim? Why might that not work?
* What other limitations might there be in this analysis?